---
title: "Data Cleaning"
author: "Darish Sakeesing"
date: "1/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Brief Introduction

This project is an attempt at the analysis of the market of Electric Vehicles, for example, what type of electric vehicle is more popular and what manufacturers are leading the industry. The purpose of the project is to inform its audience, whether consumers or policy makers, so that they can make the best decisions in their situation. 

Before we can use the data, we have to import it and clean it.

```{r Importing}
df_2012 <- read.csv('data/ev_2012.csv', header = T)
df_2013 <- read.csv('data/ev_2013.csv', header = T)
df_2014 <- read.csv('data/ev_2014.csv', header = T)
df_2015 <- read.csv('data/ev_2015.csv', header = T)
df_2016 <- read.csv('data/ev_2016.csv', header = T)
df_2017 <- read.csv('data/ev_2017.csv', header = T)
df_2018 <- read.csv('data/ev_2018.csv', header = T)
df_2018 <- read.csv('data/ev_2019.csv', header = T)

cumulative_df <- read.csv('data/cumulative_sales.csv', header = T)

```

```{r Cleaning}

```
